---
title: "Assignment 2 – Data Analysis using R Programming"
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancyhead[CO,CE]{Assignment 2 – Data Analysis using R Programming}
- \fancyfoot[CO,CE]{github - https://github.com/Chippybaby19/R-project}
- \fancyfoot[LE,RO]{\thepage}


output: pdf_document
date: '2022-04-08'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Datasets
Using UScereal Dataset from the library Mass

```{r UScereal}
library(MASS)
??UScereal
```
## R Datasets
Creating a new dataframe

```{r UScereal1}
UScereal1 <- UScereal
```

Printing the Structure of the dataset
```{r UScereal1_}
str(UScereal1)
```

Listing the variables inside the dataset
```{r Usc_2}
names(UScereal1)
```

Printing top 15 rows of the dataset
```{r Usc_3}
head(UScereal1, 15)
```

User defined function for one variable Finding the average of any given variable
```{r Usc_4}
ave<-function(x){ mn<-mean(x)
return(mn) }
ave(UScereal1$potassium)
```

Use data manipulation techniques and filter rows based on any logical criteria that exist in your dataset. Filtering for any manufacturer
```{r Usc_5}
manu1 <- UScereal1[UScereal1$mfr=='K',]
head(manu1)
```

Identify the dependent & independent variables and use reshaping techniques and create a new data frame by joining those variables from your dataset.
```{r Usc_6}
library(tidyr)
UScereal_long <- gather(UScereal1, market_info, value, calories:vitamins, factor_key=TRUE)
```

```{r Usc_7}
UScereal_long
```

Remove missingvalues in your dataset.
```{r Usc_8}
 library(ggplot2)
remove_missing(UScereal1, na.rm = FALSE, vars = names(UScereal1), name = "", finite = FALSE)
```

Identify and remove duplicated data in your dataset
```{r Usc_9}
UScereal1[!duplicated(UScereal1), ]
```

Reorder multiple rows in descending order
```{r Usc_10}
library(dplyr)
UScereal1 %>% arrange(mfr, shelf)
```

Rename some of the column names in your dataset
```{r Usc_11}
names(UScereal1)[names(UScereal1)=="calories"] <- "calorie_content"
names(UScereal1)[names(UScereal1)=="mfr"] <- "Manufacturer"
head(UScereal1)
```

Add new variables in your data frame by using a mathematical function (for e.g. – multiply an existing column by 2 and add it as a new variable to your data frame)

```{r Usc_12}
#Finding Kilocalories
UScereal1 %>%
  mutate(Kilocalories = calorie_content/1000)
```

Create a training set using random number generator engine.
```{r Usc_13}
 RandomNum <- runif(50, 1, 250)
RandomNum
```


Print the summary statistics of your dataset
```{r Usc_14}
summary(UScereal1)
```

Use any of the numerical variables from the dataset and perform the following statistical functions • Mean • Median • Mode • Range
```{r Usc_15}
 #Mean function
calorie_mean<-mean(UScereal1$calorie_content, na.rm=TRUE)
print(calorie_mean)
```

```{r Usc_16}
#Median function
calorie_median<-median(UScereal1$calorie_content, na.rm=TRUE)
print(calorie_median)
```

```{r Usc_17}
# To find Mode we need to create a function.
findmode <- function(x) {
mode_value <- unique(x) 
mode_value[which.max(tabulate(match(x, mode_value)))]
## Calculate the mode using the function.
calorie_mode <- findmode(UScereal1$calorie_content)
print(calorie_mode)
}
```

```{r Usc_18}
 #Mean function
 #To find Range
calorie_range <- max(UScereal1$calorie_content, na.rm=TRUE) - min(UScereal1$calorie_content, na.rm=TRUE)
print(calorie_range)
```

## Scatter Plot
Plot a scatter plot for any 2 variables in your dataset
```{r Usc_19}
 #Mean function
 plot(UScereal1$sugars, UScereal1$calorie_content, col='red', pch=19,
     xlab='Sugars', ylab='Calories', main='Scatterplot of Two Variables')
```

## Bar Plot
bar plot for any 2 variables in your dataset
```{r Usc_20}
# Add labels to bar plot
barplot(UScereal1$calorie_content,
names.arg = UScereal1$Manufacturer,
main = "Bar Plot of 2 variables",
xlab = "Manufacturer",
ylab = 'Calories')
```


## Linear Regression Model
Find the correlation between any 2 variables by applying least square linear regression model
```{r Usc_21}
# Linear Regression Model
model1 <- lm(UScereal1$calorie_content~UScereal1$sugars)
yfit<-model1$fitted.values
plot(UScereal1$sugars, UScereal1$calorie_content, xlab='Sugar', ylab='Calories', 
main='Calorie Content based on Sugars')
lines(UScereal1$sugars, yfit)
```

## Residual Plot
```{r Usc_22}
#residual plot
residuals <- UScereal1$calorie_content - yfit
plot(UScereal1$sugars, residuals, xlab='Sugar', ylab='Residuals', main='Residual Plot')
```

```{r Usc_23}
#Correlation
cor(UScereal1$sugars, UScereal1$calorie_content)
```



